```{r}
library("TCGAbiolinks")
library("survival")
library("survminer")
```

1- Load the data
```{r}
tcga_data = readRDS(file = "tcga_data.RDS")
```

2- Extract clinical data
```{r}
# extract clinical data
clinical = tcga_data@colData

dim(clinical)
```
3- Print column names
```{r}
colnames(clinical)
```

4- Need to extract the race variable, and a few more besides
```{r}
# we are only interested in the "Primary solid Tumor" cases for survival
clin_df = clinical[clinical$definition == "Primary solid Tumor",
                    c("patient",
                      "vital_status",
                      "days_to_death",
                      "days_to_last_follow_up",
                      "race")]
```

5- we need to change part of this information in a way that is acceptable to the methods from the `survival` package we are using:

```{r}
# create a new boolean variable that has TRUE for dead patients
# and FALSE for live patients
clin_df$deceased = clin_df$vital_status == "Dead"

# create an "overall survival" variable that is equal to days_to_death
# for dead patients, and to days_to_last_follow_up for patients who
# are still alive
clin_df$overall_survival = ifelse(clin_df$deceased,
                                   clin_df$days_to_death,
                                   clin_df$days_to_last_follow_up)

# show first 10 samples
head(clin_df)
```

6- let's look at race column
```{r}
clin_df$race
```
7- find the unique races
```{r}
unique(clin_df$race)
```
We see there are 5 categories among which we are interested in "asian","white","black or african american" and stack "american indian or alaska native" and "not reported" into a new group of "others"

8- Create new column as "new_race"

```{r}
clin_df$new_race = clin_df$race

# we replace those with race "not reported" to "others"
clin_df[which(clin_df$new_race == "not reported"), "new_race"] = "others"

# we replace those with race "american indian or alaska native" to "others"
clin_df[which(clin_df$new_race == "american indian or alaska native"), "new_race"] = "others"
```


9- To make sure if we have done our job well
```{r}
unique(clin_df$new_race)
```

10- We can now fit the survival model with the new_race groups:

```{r}
fit = survfit(Surv(overall_survival, deceased) ~ new_race, data=clin_df)

# we can extract the survival p-value and print it
pval = surv_pvalue(fit, data=clin_df)$pval
print(pval)
```

```{r}
# we produce a Kaplan-Meier plot from the fitted model
ggsurvplot(fit, data=clin_df, pval=T, risk.table=T, risk.table.height=0.35)
```

Null hypothesis: there is no difference between the race survival curves (i.e. the probability of an event occurring at any time point is the same for each population)

This gives a P value of 0.34 which indicates no significant difference between the population survival curves or there is no correlation between race and survival. So the null hypothesis is not rejected

