---
title: "Understanding P values"
author: "Carl de Boer"
date: "05/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
```


# A simulation of coin flips

In this simulation, we are simulating `numFlips` coin flips from `numCoins` unbiased coins (50/50 odds for heads). We are then calculating the P-value for how biased each coin is towards heads. Play with this simulation. 

What happens as you change the number of flips? What happens as you change the number of coins?

```{r}
numFlips = 10;
numCoins = 10000;
# rbinom returns random numbers generated from the binomial distribution (e.g. coin flips), with the probability of heads=0.5
coinFlips = data.frame(heads = rbinom(numCoins, numFlips, prob = 0.5))
#pbinom returns the p-value of observing the given number of successes (e.g. heads) given however many trials (e.g. flips), and the null hypothesis probability of success (0.5 for unbiased)
coinFlips$P = pbinom(coinFlips$heads, numFlips, prob=0.5)
```

Make a histogram of the P values
```{r}
p =  ggplot(coinFlips, aes(x=P)) + geom_histogram(binwidth = 0.01) ; print(p)
```

Make a cumulative distribution function of the P-values
```{r}
p =  ggplot(coinFlips, aes(x=P)) + stat_ecdf() + geom_abline(slope=1, intercept = 0, colour="red"); print(p)

```

# Min P-values

In this simulation, we are flipping many coins many times. 

```{r}
numFlips = 1000;
numCoins = 1000000;
coinFlips = data.frame(heads = rbinom(numCoins, numFlips, 0.5))
coinFlips$P = pbinom(coinFlips$heads, numFlips, prob=0.5)
```

What is the min P-value we obtained for a coin?
```{r}
min(coinFlips$P)
```

Given this P-value, was this particular coin biased?

# Multiple hypothesis testing

Now we're going to explore multiple hypthesis testing and correction. Here, we are using Benjamini-Hochberg FDR correction. In this case, the corrected P values (Q values) correspond to the false discovery rate if you call that test "significant". That is to say, if you call coin k biased, the approximately Q[k] of coins where the P-value was less than P[k] are not really biased.

We have two sets of coin flips here. One where the coin is biased, the other where it is unbiased. We'll see what happens when we correct their P-values together or separately.

```{r}
numFlips = 1000;
numCoins = 1000000;
coinFlipsUnbiased = data.frame(heads = rbinom(numCoins, numFlips, 0.5))
coinFlipsUnbiased$P = pbinom(coinFlipsUnbiased$heads, numFlips, prob=0.5)
coinFlipsUnbiased$Q = p.adjust(coinFlipsUnbiased$P, method="BH")
coinFlipsUnbiased$isbiased="unbiased"

#Coin flips for a biased coin

coinFlipsBiased = data.frame(heads = rbinom(numCoins, numFlips, 0.47))
coinFlipsBiased$P = pbinom(coinFlipsBiased$heads, numFlips, prob=0.5)
coinFlipsBiased$Q = p.adjust(coinFlipsBiased$P, method="BH")
coinFlipsBiased$isbiased="biased"

coinFlipsBoth =rbind(coinFlipsUnbiased,coinFlipsBiased)
coinFlipsBoth$Q = p.adjust(coinFlipsBoth$P, method="BH") # correcting P-values together with B-H FDR correction

```

## First let's look at the unbiased coin

Make a histogram of the Q values
```{r}
p =  ggplot(coinFlipsUnbiased, aes(x=Q)) + geom_histogram(binwidth = 0.01) ; print(p)
```

Make a cumulative distribution function of the Q values
```{r}
p =  ggplot(coinFlipsUnbiased, aes(x=Q)) + stat_ecdf() + geom_abline(slope=1, intercept = 0, colour="red"); print(p)

```

What fraction of the unbiased coins are significant at an FDR of 12.5%?

## Now let's look at the biased coin

Make a histogram of the Q values
```{r}
p =  ggplot(coinFlipsBiased, aes(x=Q)) + geom_histogram(binwidth = 0.01) ; print(p)
```

Make a cumulative distribution function of the Q values
```{r}
p =  ggplot(coinFlipsBiased, aes(x=Q)) + stat_ecdf() + geom_abline(slope=1, intercept = 0, colour="red"); print(p)

```

What fraction of the biased coins are significant at an FDR of 12.5%?

## Now let's look at the effect of correcting the biased and unbiased coins together

Make a histogram of the Q values
```{r}
p =  ggplot(coinFlipsBoth, aes(x=Q)) + geom_histogram(binwidth = 0.01) + facet_grid(isbiased ~ .) ; print(p)
```

Make a cumulative distribution function of the Q values
```{r}
p =  ggplot(coinFlipsBoth, aes(x=Q, colour=isbiased)) + stat_ecdf() + geom_abline(slope=1, intercept = 0, colour="red"); print(p)

```

What fraction of the biased coins are significant at an FDR of 12.5%?

What fraction of the unbiased coins are significant at an FDR of 12.5%?

Why is this different from controlling for the FDR separately for the biased and unbiased coins?
(consider what the FDR means)


